---
title: "Practical Machine Learning Final Project"
author: "Sang Cho"
date: "July 19, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Summary

The purpose of this project is to identify the labels in the test dataset by building a model based on the qualities of the activity the participant.

# Data Prep

Download and load the train and test datasets.  Also, within the training dataset, I've splitted the validation set in order to test the accuracy of the model.


```{r}
download.file('https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv',destfile = "training.csv")
training <- read.csv("training.csv",na.strings=c("NA",""))

download.file('https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv',destfile = "test.csv")
test <- read.csv("test.csv",na.strings=c("NA",""))
```

# Data Cleaning

Remove columns with (1. 90% or more NA values  2.  First 6 ID columns not used for the model)
For remaining columns, impute the missing values with the average

```{r}
training <- training[,-(1:6)]
training <- training[,colMeans(is.na(training)) < 0.1]


imp <- function(x) {
  for(i in 1:ncol(x))
  {
    x[is.na(x[,i]), i] <- mean(x[,i], na.rm = TRUE)
  }
}

imp(training)
```

# Training / Validation Datasets

Within the training dataset, I've splitted the validation set in order to test the accuracy of the model.

```{r}
library(caret)
partition <- createDataPartition(training$classe,p=0.7,list=FALSE)
train.raw <- training[partition,]
test.raw <- training[-partition,]
```

# CART Model


```{r}
set.seed(111)

Control <- trainControl(method='cv', number = 5)
cartmod <- train(classe ~ ., data = train.raw,method="rpart",trControl=Control)
cartpred <- predict(cartmod,test.raw)
confusionMatrix(cartpred,test.raw$classe)
```

CART model was built with 5-fold CV to select the optimal tuning parameters.
However, the accuracy rate  ~ 50% called for a different model.


# Random Forest


```{r}
rfmod <- train(classe ~ ., data = train.raw,method="rf",trControl=Control)
rfpred <- predict(rfmod,test.raw)
confusionMatrix(rfpred,test.raw$classe)
```

Random Forest model was built with 5-fold CV to select the optimal tuning parameters.
The accuracy rate was satisfactory and I have chosen the Random Forest model to proceed.
On the side note, I've also built another model with gbm; however the accuracy rate was lower than the RF and the run time was extremely long so I have not included the run in the markdown.





